---
title: "Day62: Data exploration"
author: "csiu"
date: "April 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(pheatmap)
```

```{r message=FALSE}
# Load data
train <- read_csv("../input/train.csv")

num_samples <- nrow(train)
num_columns <- ncol(train)
paste("Number of samples:", num_samples, "&",
      "Number of columns:", num_columns)
```

## How complete is the data?

```{r}
(stat_missing_data <-
  train %>%
  summarise_each_(funs(is.na(.) %>% sum),
                  vars = paste0("`", colnames(train), "`")) %>%
  gather(column_name, num_na) %>%
  mutate(per_na = num_na/num_samples) %>%
  filter(num_na != 0)
) %>%

  ggplot(aes(x = reorder(column_name, -num_na),
             y = per_na)) +
  geom_bar(stat="identity") +
  ggtitle("Missing data") +
  coord_cartesian(ylim = c(0, 0.5)) +
  ylab("Proportion of NA samples") +
  xlab("Features") +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
```

```{r}
stat_missing_data %>%
  {nrow(.) / num_columns} %>%
  {sprintf("Proportion of columns containing at least 1 NA: %.2f", .)}
```

### What columns are there?

There are a lot of columns -- I wonder what they are. Let me try grouping the columns by name

```{r}
stat_class_type <-
  lapply(colnames(train), function(x){class(train[[x]])}) %>%
  unlist() %>%
  {data.frame(
    column_name = colnames(train),
    class_type = .
  )} %>%
  arrange(class_type, column_name)
```

```{r}
remove_column <- function(all_features, features_to_rm){
  paste0("^", features_to_rm, "$") %>%
  paste0(collapse = "|") %>%
  grep(all_features, value = TRUE, invert = TRUE)
}


train_columns <- colnames(train) %>%
  sort()

# write_lines(train_columns, "day63_features.txt")

features <- NULL
features$sex <- 
  grep("_female|_male|_all", train_columns, value = TRUE) %>%
  {grep("full_all", ., value = TRUE, invert = TRUE)}
train_columns <- remove_column(train_columns, features$sex)

features$distance <-
  grep("_km", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$distance)

features$avg_price <-
  grep("avg", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$avg_price)

features$raion <-
  grep("_raion$", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$raion)

features$count <-
  grep("_count_", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$count)

features$part <-
  grep("part", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$part)

features$line1 <-
  grep("1line", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$line1)

features$sqm <-
  grep("sqm", train_columns, value = TRUE)
train_columns <- remove_column(train_columns, features$sqm)


## The other columns
lapply(train_columns, function(x){class(train[[x]])}) %>%
  unlist() %>%
  {data.frame(
    column_name = train_columns,
    class_type = .
  )} %>%
  arrange(class_type, column_name)

features$other <- train_columns
train_columns <- remove_column(train_columns, features$other)

# train %>% select_(.dots = train_columns)
```

```{r results='asis'}
for (name in names(features)){
  cat(paste0("**", name, "**"))
  cat("\n")
  cat(features[[name]])
  cat("\n\n")
}
```

### Floor space

I also wanted to look at full_sq and life_sq.

```{r}
train_sq <-
  train %>%
  select(id, ends_with("sq"), price_doc)

head(train_sq)
```

```{r}
sum(is.na(train_sq$kitch_sq)) %>%
  {. / nrow(train_sq) * 100} %>%
  {sprintf("Number of NAs in kitch_sq: %0.1f%%", .)}
```


```{r}
train_sq %>%
  ggplot(aes(x = full_sq, y = life_sq)) +
  geom_point()
```

Outliers?

```{r}
train_sq_tidy <-
  train_sq %>%
  mutate(diff = full_sq - life_sq)

## Error?
filter(train_sq_tidy,
       abs(diff) > 2000)
```

```{r}
train_sq %>%
  mutate(price_doc = price_doc/1e6) %>%
  
  ggplot(aes(x = full_sq, y = price_doc)) +
  geom_point() +
  scale_x_log10() +
  ylab("Price per million")
```

```{r}
train_sq %>%
  mutate(price_doc = price_doc/1e6) %>%
  
  ggplot(aes(x = life_sq, y = price_doc)) +
  geom_point() +
  scale_x_log10() +
  ylab("Price per million")
```

From the plots, it does not look like the outlier points (with greater floor space) have higher prices.

```{r}
train_sq_tidy <-
  train_sq_tidy %>%
  filter(!id %in% c("3530", "13549"))
```

```{r}
train_sq_tidy %>%
  ggplot(aes(x = full_sq, y = life_sq)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red")
```

As expected, full\_sq is bigger than life\_sq

```{r}
train_sq_tidy %>%
  mutate(price_doc = price_doc/1e6) %>%
  
  ggplot(aes(x = full_sq, y = life_sq, color=price_doc)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  scale_color_gradient(low = "grey88", high = "red")
```

```{r}
train_sq_tidy %>%
  ggplot(aes(x = diff)) +
  geom_density() +
  coord_cartesian(xlim = c(-10, 100))
```

The difference between life and full sq is about 22 ft.

### Distance

The distance group seems interesting

```{r}
train_distance <-
  train %>%
  select_(.dots = features$distance)
```

```{r}
cor(train_distance, use = "complete.obs") %>%
  as.matrix() %>%
  pheatmap()
```

Water, green zone, industrial, cemetery, big road seems to be far from everything else.
