---
title: "Day37: Exploring pesticide residues in food"
output:
  html_document:
    keep_md: yes
---

I want to take a look at pesticides today. The data I use comes from Kaggle's [Pesticide Data Program (2015)](https://www.kaggle.com/usdeptofag/pesticide-data-program-2015/kernels) data set uploaded by the United States Department of Agriculture.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r load-libs, message=FALSE}
library(dplyr)
library(ggplot2)
library(DBI)
```

```{r load-data, results='hide'}
con <- dbConnect(RSQLite::SQLite(), dbname="data/database.sqlite")
results <- dbGetQuery(con, 'SELECT * FROM resultsdata15')
dbDisconnect(con)
```

```{r load-codes, message=FALSE, warning=FALSE}
code <- NULL
code$commod <-
  readr::read_csv("data/commodity_codes.csv",
                  col_names = c("code", "label", "sample_size"),
                  skip = 1)
```

### Glimpse of the data

Size of the data (rows, columns):

```{r}
dim(results)
```

```{r}
head(results) %>%
  knitr::kable()
```

### Dealing with missing data

```{r process-results}
# Replace empty string with NA
results[results==""] <- NA
```

```{r deal-with-na}
# How many NAs?
(na_in_column <-
  colSums(is.na(results))) %>%
  sort(decreasing = TRUE) %>%
  as.data.frame()

## Remove columns with NAs
results <- results[na_in_column==0]
```

```{r}
# Tidy data
results <-
  results %>%
  mutate(determin = gsub("[\r\n]", "", determin)) %>%
  tbl_df()
```

### The columns

- sample_pk
- commod - COMMODITY TYPE in 2015 PDP Samples
- commtype - COMMODITIES Sampled/Analyzed by PDP in 2015 (Fresh Product Unless Otherwise Noted)
- lab - PDP Participating LABORATORIES in 2015
- pestcode - Compounds (PESTICIDES) Analyzed by PDP
- testclass - TEST (COMPOUND) CLASS in 2015 PDP Analytical Results
- lod - Limit of Detection
- conunit - Concentration/LOD Unit-of-Measure Code
- mean - MEAN RESULT in 2014 PDP Analytical Results (O, A, and R indicated Positive Detections)
- extract - EXTRACTION METHOD in 2015 PDP Analytical Results
- determin - DETERMINATIVE METHOD in 2015 PDP Analytical Results

### A simple question

Which commodity type has the highest level of pesticides?

```{r box-commtype}
# Check if the units are all the same
results$conunit %>% unique()

results %>%
  ggplot(aes(x = commtype, y = lod)) +
  geom_boxplot()
```

FR = Fresh, FZ = Frozen, OT = Other

```{r box-commod}
select(code$commod, code, label) %>%
  right_join(results, by = c("code"="commod")) %>%
  mutate(label = ifelse(is.na(label), code, label)) %>%
  plyr::rename(c("label" = "commod")) %>%

  ggplot(aes(x = reorder(commod, lod, FUN=median),
             y = lod, fill = commtype)) +
  geom_boxplot() +
  xlab("Commodities") +
  ylab("Limit of Detection in Parts-per-Million") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```
